---
title: "Econ 204: HW 2"
author: "Sydney Mayes, Olivia Somhegyi, Rachel Spellenberg"
date: "2023-04-29"
output:  
  html_document:
    theme: simplex
    code_folding: hide
---

```{r setup,  include = TRUE, echo=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, results = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(equatiomatic)
```


```{r}

# In this homework you will be simulating alternative approaches for reducing carbon emissions among four UC campuses (UCSB, UCLA, UCI, and UCD). The dataset for this assignment is available in the file “CarbonUC.csv”, available on Canvas. The dataset contains estimates of the marginal cost of reducing carbon emission for each campus for different levels of abatement (in tons of carbon). A few notes:
# ● Current (i.e. baseline) emission for each campus are: UCSB = 90,000 tons, UCLA = 250,000 tons, UCI = 100,000 tons, UCD = 160,000 tons.
# ● Prices are in $/ton and quantities are in tons of CO2.
# ● Before opening R, I recommend drawing graphs by hand to make sure you know what you want to do

# Clean and format dataset
carbon_uc <- read_csv(here('data', 'CarbonUC.csv')) %>% 
  clean_names()

long_uc <- pivot_longer(carbon_uc, cols = ucsb:ucd, names_to = 'uc', values_to = 'marginal_cost')

```


#### Question 1
For each campus, plot the marginal abatement cost data and estimate a linear regression model with
zero intercept. These estimated linear regressions will be your estimates of the marginal abatement
cost curve for each campus, which you will use for the remainder of this analysis. Write the equation
for the marginal abatement cost curve for each campus.

```{r}
# For each campus, plot the marginal abatement cost data and estimate a linear regression model with zero intercept.1 These estimated linear regressions will be your estimates of the marginal abatement cost curve for each campus, which you will use for the remainder of this analysis. Write the equation for the marginal abatement cost curve for each campus.


# Estimating linear regression models with zero intercept for each UC
# UCSB
ucsb_model <- lm(marginal_cost ~ 0 + abatement, data = filter(long_uc, uc == 'ucsb'))
#p = ucsb_model$coefficients*q
#extract_eq(ucsb_model, use_coefs = TRUE, coef_digits = 5)
# p = .001*q
ucsb_slope <- ucsb_model$coefficients
  
# UCLA
ucla_model <- lm(marginal_cost ~ 0 + abatement, data = filter(long_uc, uc == 'ucla'))
# ucla_model$coefficients
# p = .0007*q
ucla_slope <- ucla_model$coefficients

# UCI
uci_model <- lm(marginal_cost ~ 0 + abatement, data = filter(long_uc, uc == 'uci'))
# uci_model$coefficients
# p = .002 *q
uci_slope <- uci_model$coefficients

# UCD
ucd_model <- lm(marginal_cost ~ 0 + abatement, data = filter(long_uc, uc == 'ucd'))
# ucd_model$coefficients
# p = .003 *q
ucd_slope <- ucd_model$coefficients


# Plotting the marginal abatement costs data for each school
ggplot(data = long_uc, aes(x = abatement, y = marginal_cost, col = uc)) +
  geom_point() +
  geom_abline(slope = ucsb_slope) +
  geom_abline(slope = ucla_slope) +
  geom_abline(slope = uci_slope) +
  geom_abline(slope = ucd_slope) +
  #facet_wrap(~uc, ncol = 2) +
  scale_color_viridis_d() +
  labs(x = 'Tons of Carbon Abated', y = 'Price in dollars/ton of Carbon' ) +
  theme_bw()

```

##### UCSB: `r equatiomatic::extract_eq(ucsb_model, wrap = TRUE, use_coefs = TRUE, coef_digits = 5)`
##### UCLA:  `r equatiomatic::extract_eq(ucla_model, wrap = TRUE, use_coefs = TRUE, coef_digits = 5)`
##### UCI:  `r equatiomatic::extract_eq(uci_model, wrap = TRUE, use_coefs = TRUE, coef_digits = 5)`
##### UCD:  `r equatiomatic::extract_eq(ucd_model, wrap = TRUE, use_coefs = TRUE, coef_digits = 5)`


#### Question 2

Using your estimated marginal abatement cost curves, together with the baseline emissions, derive each campus’ demand curve for carbon. In other words, how much would each campus be willing to pay to emit the first ton of carbon, the second ton of carbon, etc.2 Plot each campus’ demand curve for emissions on a single graph. Which campus is willing to pay the most for the first ton of carbon emissions?

```{r}

# UCSB = 90,000 tons, UCLA = 250,000 tons, UCI = 100,000 tons, UCD = 160,000 tons.
# A = Qbar - Q

ucsb_base = 90000
ucla_base = 250000
uci_base = 100000
ucd_base = 160000

# e = amount of tons emitted

# UCSB
ucsb_price <- function(e) {
   p = ucsb_slope*(ucsb_base - e)
   return(p)
}

# UCLA
ucla_price <- function(e) {
   p = ucla_slope*(ucla_base - e)
   return(p)
}

# UCI
uci_price <- function(e) {
   p = uci_slope*(uci_base - e)
   return(p)
}

# UCD
ucd_price <- function(e) {
   p = ucd_slope*(ucd_base - e)
   return(p)
 }

demand_df <- data.frame(e = seq(0, 250000, length.out = 100)) %>% 
  mutate(ucsb_cost = pmax(0, ucsb_price(e))) %>% 
  mutate(ucla_cost = pmax(0, ucla_price(e))) %>% 
  mutate(uci_cost = pmax(0, uci_price(e))) %>% 
  mutate(ucd_cost = pmax(0, ucd_price(e)))

# e = x-axis; y = ucsb_cost

demand_df_long <- pivot_longer(demand_df, cols = ucsb_cost:ucd_cost, names_to = 'campus', values_to = 'marginal_wtp')

gg_demand <- ggplot(data = demand_df_long) +
  geom_line(aes(x = e, y = marginal_wtp, col = campus)) +
  scale_color_viridis_d() +
  labs(x = 'Tons of Carbon Emitted', y = 'Price in dollars/ton of Carbon' ) +
  theme_bw()
  
gg_demand
  

```
Based on the graph, UCD would pay the most for the first ton emitted.

#### Question 3
Here you will analyze three alternative interventions for reducing carbon emissions across these four campuses. Please answer the questions about each intervention listed below.


a. Intervention #1: Carbon Tax: Each campus must pay a tax of $50/ton for each ton of emissions. How much would each campus choose to emit? How much would each campus abate? What is the total amount of emissions across all four campuses? How much tax revenue is generated?
```{r}
# a = qbar - q
# q = qbar - a
tax = 50
ucsb_emissions = ucsb_base - (tax/ucsb_slope)

ucla_emissions = ucla_base - (tax/ucla_slope)

uci_emissions = uci_base - (tax/uci_slope)

ucd_emissions = ucd_base - (tax/ucd_slope)

ucsb_emissions


ucsb_abate = tax/ucsb_slope
ucla_abate = tax/ucla_slope
uci_abate = tax/uci_slope
ucd_abate = tax/ucd_slope

total_emissions = ucsb_emissions + ucla_emissions + uci_emissions + ucd_emissions

ucsb_revenue = tax*ucsb_emissions 
ucla_revenue = tax*ucla_emissions 
uci_revenue = tax*uci_emissions
ucd_revenue = tax*ucd_emissions

total_revenue = ucsb_revenue + ucla_revenue + uci_revenue + ucd_revenue
```

UCSB would choose to emit 40209.19 tons of carbon and would abate 49790.81 
UCLA would choose to emit 178717.5 tons of carbon and would abate 71282.47 
UCI would choose to emit 75096.24 tons of carbon and would abate  24903.76 
UCI would choose to emit 143466.4 tons of carbon and would abate  16533.62 

Total emissions:  437489.3 tons of carbon

Tax revenue for UCSB:  2010460 dollars
Tax revenue for UCLA:  8935877 dollars
Tax revenue for UCI:  3754812  dollars
Tax revenue for UCD:  7173319  dollars
Total Tax Revenue =  21874467 dollars



b. Intervention #2: Carbon Ban: Carbon emissions are banned at each campus. What is the total cost of this ban to each campus?
```{r}
ucsb_ban = 0.5 * ucsb_base * (ucsb_slope*ucsb_base) # 4067016 
ucla_ban = 0.5 * ucla_base * (ucla_slope*ucla_base) # 21919836 
uci_ban = 0.5 * uci_base * (uci_slope*uci_base) # 10038644 
ucd_ban = 0.5 * ucd_base * (ucd_slope*ucd_base) # 38709018 

```

Total cost of the ban for UCSB is 4067016 dollars
Total cost of the ban for UCLA is 21919836 dollars
Total cost of the ban for UCI is 10038644 dollars
Total cost of the ban for UCD is 38709018 dollars



c. Intervention #3: Cap without Trade: Each campus is capped at 100,000 tons of carbon. How much abatement would be required by each campus to comply with this cap? What is the marginal cost of the last unit of abatement for each campus? What is the total cost of abatement for each campus to comply with this new cap?

```{r}
cap = 100000
ucsb_cap_notrade = max(0, ucsb_base - cap) # 0
ucla_cap_notrade = max(0, ucla_base - cap) # 150000
uci_cap_notrade = max(0, uci_base - cap) # 0
ucd_cap_notrade = max(0, ucd_base - cap) # 60000

```

UCSB doesn't need to abate anything
UCLA must abate 150000 tons of carbon
UCI doesn't need to abate anything
UCD must abate 60000 tons of carbon


What is the marginal cost of the last unit of abatement for each campus?
```{r}
ucla_cap_notrade_marginal <- ucla_price(150000) #  70.14347 
ucd_cap_notrade_marginal <- ucd_price(60000) # 302.4142 

```

Marginal cost of the last unit of abatement for UCLA is 70.14347 dollars
Marginal cost of the last unit of abatement for UCLA is 302.4142 dollars
UCSB and UCI were under the cap.

What is the total cost of abatement for each campus to comply with this new cap?
```{r}
ucla_cap_notrade_total <- 0.5 * ucla_cap_notrade_marginal * (250000-100000) # 5260761 
ucd_cap_notrade_total <- 0.5 * ucd_cap_notrade_marginal * (160000-100000) #  9072426
```

The total cost of abatement for UCLA is 5260761 dollars.
The total cost of abatement for UCLA is 9072426 dollars.
(No cost for the other campuses)

d. Intervention #4: Cap with Trade: Suppose each campus is capped at 100,000 tons of carbon, but trade is allowed. With trade, total emissions will be 400,000 but some campuses may emit more or less than 100,000. How much would each campus end up emitting after trade? What is the equilibrium trading price of carbon permits? For each campus, calculate whether they are a buyer or a seller of permits.

```{r}

demand_df_2 <- data.frame(p = seq(0, 500, length.out = 100)) %>% 
  mutate(ucsb_e = pmax(0, ucsb_base-(p/ucsb_slope))) %>% 
  mutate(ucla_e = pmax(0, ucla_base-(p/ucla_slope))) %>% 
  mutate(uci_e = pmax(0, uci_base-(p/uci_slope))) %>% 
  mutate(ucd_e = pmax(0, ucd_base-(p/ucd_slope))) %>% 
  mutate(uc_agg_e = ucsb_e + ucla_e + uci_e + ucd_e)

p_star <- spline(x = demand_df_2$uc_agg_e, y = demand_df_2$p, xout = 400000)$y

demand_df_eq <- data.frame(p = 61.5345) %>% 
  mutate(ucsb_e = pmax(0, ucsb_base-(p/ucsb_slope))) %>% 
  mutate(ucla_e = pmax(0, ucla_base-(p/ucla_slope))) %>% 
  mutate(uci_e = pmax(0, uci_base-(p/uci_slope))) %>% 
  mutate(ucd_e = pmax(0, ucd_base-(p/ucd_slope))) %>% 
  mutate(uc_agg_e = ucsb_e + ucla_e + uci_e + ucd_e)


# 100,000 - 28722.95
# 162273.4 - 100000
# 100000 - 69351.19
# 139652.2 - 100000

```

The equilibrium trading price of carbon permits is 61.5345 dollars
UCSB will emit 28722.95 tons of carbon; they will sell 71277.05 permits
UCLA will emit 162273.4	tons of carbon; they will buy 62273.4 permits
UCI will emit 69351.19 tons of carbon; they will sell 30648.81 permits
UCD will emit 139652.2 tons of carbon; they will buy 39652.2 permits

#### Question 4:


